Exploring Vectorization and unrolling with SMT

### 2022-05-23 article/article.tex
add axya to article
Nouvelles courbes

\subsection{Impact of unrolling factor on Intel CascadeLake}
\begin{figure*}[h!]
%  \begin{subfigure}[h]{0.45\textwidth}
%  \includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/icc-2021.2.0.pdf}
%  \caption{icc 2021.2.0}
%  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/icc-19.1.3.pdf}
  \caption{icc 19.1.3}
  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/pbqp.pdf}
  \caption{Clang 15 with PBQP register allocation}
  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/greedy.pdf}
  \caption{Clang 15 with greedy register allocation}
  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/basic.pdf}
  \caption{Clang 15 with basic register allocation}
  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/gcc-11.9.pdf}
  \caption{GCC 11.9}
  \end{subfigure}\hfill ~
  \caption{Block DGEMM Performance on Intel CascadeLake w.r.t. register and spill usage. Performance is given in \% of peak flops/cycle (here 32). The block (BI,BK,BJ) is 64x256x64. Each code represents a different register block size, obtained through the three unrolling factors. Codes with register block size $\leq 32$ are in blue (should fit in registers whatever the schedule), other codes are in grey (with spill) and red (without spill).  \label{fig:cascadelake}}
\end{figure*}
 For all plots in Fig.\ref{fig:cascadelake}, the same codes with the same vectorization expressed with MIPP are
compiled. All codes are tested 40 times, the best performance is shown. Only one memory allocation is used for all tests. Points shown in blue correspond to codes that should not require
more SIMD registers than the physical registers (32 here, shown by the vertical red line), whatever
the scheduling chosen by the compiler. This is not the case for codes
shown in red, some schedules may manage to use only registers while
others will require spill code. The graphs show performance w.r.t. the
cumulated register count and spill used by the code. For instance, a value of 40 registers and
spill used indicates that the compiler was able to use all 32
registers and 8 additional spaces on the stack for spill. With the exception of
GCC, other compilers manage to allocate less than 32 registers for all
blue codes. The max performance on this machine is normally 32 flops/cycles (with 2 FMAs operating on AVX512 vectors).

@@

\begin{figure*}[h!]
% For clang 15
{    \scriptsize
    \begin{tabular}{|c|c|c|c|}
      \hline
      \multicolumn{4}{|c|}{Code performance with Clang} \\
      \hline
      Unrolling & flop/cycle & block  & ILP\\
      ixkxj& in peak \% & size & \\
      \hline
14x1x2 & 93.59 & 44 & 28 \\
4x1x6 & 93.45 & 34 & 24 \\
5x3x4 & 93.34 & 47 & 20 \\
6x1x4 & 93.78 & 34 & 24 \\
6x2x3 & 93.84 & 36 & 18 \\
6x2x4 & 93.95 & 44 & 24 \\
5x1x5 & 94.79 & 35 & 25 \\
6x3x3 & 94.47 & 45 & 18 \\
6x3x4 & 94.53 & 54 & 24 \\
6x4x3 & 94.05 & 54 & 18 \\
7x2x3 & 94.81 & 41 & 21 \\
8x1x3 & 94.65 & 35 & 24 \\
7x3x3 & 95.05 & 51 & 21 \\
8x2x3 & 95.12 & 46 & 24 \\
\hline
\end{tabular}
% For icc
    \begin{tabular}{|c|c|c|c|}
      \hline
      \multicolumn{4}{|c|}{Code performance with icc 19.1.3} \\
      \hline
      Unrolling & flop/cycle & block  & ILP\\
      ixkxj& in peak \% & size & \\
      \hline
4x3x3 & 78.09 & 33 & 12 \\
3x2x6 & 79.18 & 36 & 18 \\
3x2x8 & 79.61 & 46 & 24 \\
3x1x8 & 81.90 & 35 & 24 \\
4x3x4 & 81.55 & 40 & 16 \\
7x2x2 & 83.77 & 32 & 14 \\
4x2x4 & 84.94 & 32 & 16 \\
6x2x3 & 84.73 & 36 & 18 \\
5x2x4 & 87.42 & 38 & 20 \\
4x2x5 & 88.23 & 38 & 20 \\
4x1x6 & 89.93 & 34 & 24 \\
10x1x2 & 90.70 & 32 & 20 \\
5x1x5 & 93.48 & 35 & 25 \\
6x1x4 & 95.23 & 34 & 24 \\
\hline
    \end{tabular}
% For gcc
    \begin{tabular}{|c|c|c|c|}
      \hline
      \multicolumn{4}{|c|}{Code performance with gcc 11.9} \\
      \hline
      Unrolling & flop/cycle & block  & ILP\\
      ixkxj& in peak \% & size & \\
      \hline
2x5x4 & 54.09 & 38 & 8 \\
3x6x2 & 54.83 & 36 & 6 \\
2x3x5 & 55.79 & 31 & 10 \\
2x2x6 & 57.52 & 28 & 12 \\
3x3x5 & 57.37 & 39 & 15 \\
3x5x4 & 57.08 & 47 & 12 \\
3x2x5 & 59.60 & 31 & 15 \\
3x5x3 & 59.15 & 39 & 9 \\
2x4x4 & 60.30 & 32 & 8 \\
3x4x4 & 60.33 & 40 & 12 \\
3x5x2 & 60.89 & 31 & 6 \\
3x3x4 & 62.91 & 33 & 12 \\
3x4x3 & 62.37 & 33 & 9 \\
5x3x2 & 71.70 & 31 & 10 \\
\hline
    \end{tabular}
}
\caption{Performance of the best 14 codes with Clang15,  icc 19.1.3 and gcc 11.9. Only codes using 32 registers and 0 spill are considered here\label{fig:tables}. block size corresponds to $ui*uj+ui*uk+uk*uj$. $40$ means that the code requires $40$ zmm registers if all loads are scheduled before the computation. ILP corresponds here to $ui*uj$.}
\end{figure*}

\subsection{Impact of expected ILP on Intel CascadeLake}
Figure \ref{fig:cascadelakeilp} shows the performance according to the parallelism expected from unrolling, here $ui * uj$. Indeed, $uk$ does not create additionnal ILP. From this plot, we observe that for icc and clang, the best codes result from a high ILP. On the contrary, for gcc the code without spill with the highest expected ILP only reaches $42\%$ of the max performance. The maximum performance

\begin{figure*}[h!]
%  \begin{subfigure}[h]{0.45\textwidth}
%  \includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/icc-2021.2.0.pdf}
%  \caption{icc 2021.2.0}
%  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/icc-19.1.3p.pdf}
  \caption{icc 19.1.3}
  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/greedyp.pdf}
  \caption{Clang 15}
  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/gcc-11.9p.pdf}
  \caption{GCC 11.9}
  \end{subfigure}\hfill ~
  \caption{Block DGEMM Performance used on Intel CascadeLake according to expected ILP. Performance is given in \% of peak flops/cycle (here 32). The expected ILP is the product of the unrolled factors $ui * uj$. The block (BI,BK,BJ) is 64x256x64. Each code represents a different register block size, obtained through the three unrolling factors. Codes with register block size $\leq 32$ are in blue (should fit in registers whatever the schedule), other codes are in grey (with spill) and red (without spill).\label{fig:cascadelakeilp}}
\end{figure*}

@@

The experiments show that the best performing codes have the same unrolling factors and the assembly code generated are the same. 
\begin{figure*}[h!]
  \begin{subfigure}[h]{0.45\textwidth}
  \includegraphics[width=\textwidth]{../benches/gemm/haswell-64x256x64/greedy.pdf}
  \caption{Clang 15 on Intel Haswell}
  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/zen3-64x256x64/greedy.pdf}
  \caption{Clang 15 on AMD Zen3$^*$}
  \end{subfigure}
  \begin{subfigure}[h]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/zen3-64x256x64/greedy-zen2.pdf}
  \caption{Clang 15 on AMD Zen2$^*$}
  \end{subfigure}
  \caption{Block DGEMM Performance used on AVX2 architectures. Performance is given in \% of peak flops/cycle (here 8). The block (BI,BK,BJ) is 64x128x64. Each code represent a different register block size, obtained through the three unrolling factors. Codes with register block size $\leq 32$ are in blue (fit in register whatever the schedule), others in red. (c) show correlation between the two plots. $^*$ Performance scaled down to take into account turboboost effect, active on these machines\label{fig:avx2}}
\end{figure*}

>>>>

\subsection{Impact of unrolling factor on Intel CascadeLake}
\begin{figure*}[h!]
%  \begin{subfigure}[h]{0.45\textwidth}
%  \includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/icc-2021.2.0.pdf}
%  \caption{icc 2021.2.0}
%  \end{subfigure}
  \begin{subfigure}[h!]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/icc-19.1.3.pdf}
  \caption{Intel CascadeLake, AVX-512, icc 19.1.3}
  \end{subfigure}
  \begin{subfigure}[h!]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/pbqp.pdf}
  \caption{Intel CascadeLake, AVX-512, Clang 15/PBQP}
  \end{subfigure}\\
  \begin{subfigure}[h!]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/greedy.pdf}
  \caption{Intel CascadeLake, AVX-512, Clang 15/greedy}
  \end{subfigure}
  \begin{subfigure}[h!]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/basic.pdf}
  \caption{Intel CascadeLake, AVX-512, Clang 15/basic}
  \end{subfigure}\\
  \begin{subfigure}[h!]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/gcc-11.9.pdf}
  \caption{Intel CascadeLake, AVX-512, GCC 11.9}
  \end{subfigure}
    \begin{subfigure}[h!]{0.45\textwidth}
  \includegraphics[width=\textwidth]{../benches/gemm/haswell-64x256x64/greedy.pdf}
  \caption{Intel Haswell, AVX-2, Clang 15}
  \end{subfigure}\\
  \begin{subfigure}[h!]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/zen3-64x256x64/greedy.pdf}
  \caption{AMD Zen 3, AVX-2, Clang 15$^*$}
  \end{subfigure}
  \begin{subfigure}[h!]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/zen3-64x256x64/greedy-zen2.pdf}
  \caption{AMD Zen 2, AVX-2, Clang 15$^*$}
  \end{subfigure}
  \caption{Block DGEMM Performance on AVX-512 and AVX-2 architectures w.r.t. register and spill usage. Performance is given in \% of peak flops/cycle. The block (BI,BK,BJ) is 64x256x64. Each code represents a different register block size, obtained through the three unrolling factors. Codes with register block size $\leq 32$ on AVX-512, $\leq 16$ for AVX-2,  are in blue (should fit in registers whatever the schedule), other codes are in grey (with spill) and red (without spill).  \label{fig:cascadelake}}
\end{figure*}
 For all plots in Fig.\ref{fig:cascadelake}, the same codes with the same vectorization expressed with MIPP are
compiled. All codes are tested 40 times, the best performance is shown. Only one memory allocation is used for all tests. Points shown in blue correspond to codes that should not require
more SIMD registers than the physical registers (32 here, shown by the vertical red line), whatever
the scheduling chosen by the compiler. This is not the case for codes
shown in red, some schedules may manage to use only registers while
others will require spill code. The graphs show performance w.r.t. the
cumulated register count and spill used by the code. For instance, a value of 40 registers and
spill used indicates that the compiler was able to use all 32
registers and 8 additional spaces on the stack for spill. With the exception of
GCC, other compilers manage to allocate less than 32 registers for all
blue codes. The max performance on this machine is normally 32 flops/cycles (with 2 FMAs operating on AVX512 vectors).

@@

\begin{figure*}[h!]
% For clang 15
{    \scriptsize
    \begin{tabular}{|c|c|c|c|}
      \hline
      \multicolumn{4}{|c|}{Code performance with Clang} \\
      \hline
      Unrolling & flop/cycle & block  & ILP\\
      ixkxj& in peak \% & size & \\
      \hline
14x1x2 & 93.59 & 44 & 28 \\
4x1x6 & 93.45 & 34 & 24 \\
5x3x4 & 93.34 & 47 & 20 \\
6x1x4 & 93.78 & 34 & 24 \\
6x2x3 & 93.84 & 36 & 18 \\
6x2x4 & 93.95 & 44 & 24 \\
5x1x5 & 94.79 & 35 & 25 \\
6x3x3 & 94.47 & 45 & 18 \\
6x3x4 & 94.53 & 54 & 24 \\
6x4x3 & 94.05 & 54 & 18 \\
7x2x3 & 94.81 & 41 & 21 \\
8x1x3 & 94.65 & 35 & 24 \\
7x3x3 & 95.05 & 51 & 21 \\
8x2x3 & 95.12 & 46 & 24 \\
\hline
\end{tabular}
% For icc
    \begin{tabular}{|c|c|c|c|}
      \hline
      \multicolumn{4}{|c|}{Code performance with icc 19.1.3} \\
      \hline
      Unrolling & flop/cycle & block  & ILP\\
      ixkxj& in peak \% & size & \\
      \hline
4x3x3 & 78.09 & 33 & 12 \\
3x2x6 & 79.18 & 36 & 18 \\
3x2x8 & 79.61 & 46 & 24 \\
3x1x8 & 81.90 & 35 & 24 \\
4x3x4 & 81.55 & 40 & 16 \\
7x2x2 & 83.77 & 32 & 14 \\
4x2x4 & 84.94 & 32 & 16 \\
6x2x3 & 84.73 & 36 & 18 \\
5x2x4 & 87.42 & 38 & 20 \\
4x2x5 & 88.23 & 38 & 20 \\
4x1x6 & 89.93 & 34 & 24 \\
10x1x2 & 90.70 & 32 & 20 \\
5x1x5 & 93.48 & 35 & 25 \\
6x1x4 & 95.23 & 34 & 24 \\
\hline
    \end{tabular}
% For gcc
    \begin{tabular}{|c|c|c|c|}
      \hline
      \multicolumn{4}{|c|}{Code performance with gcc 11.9} \\
      \hline
      Unrolling & flop/cycle & block  & ILP\\
      ixkxj& in peak \% & size & \\
      \hline
2x5x4 & 54.09 & 38 & 8 \\
3x6x2 & 54.83 & 36 & 6 \\
2x3x5 & 55.79 & 31 & 10 \\
2x2x6 & 57.52 & 28 & 12 \\
3x3x5 & 57.37 & 39 & 15 \\
3x5x4 & 57.08 & 47 & 12 \\
3x2x5 & 59.60 & 31 & 15 \\
3x5x3 & 59.15 & 39 & 9 \\
2x4x4 & 60.30 & 32 & 8 \\
3x4x4 & 60.33 & 40 & 12 \\
3x5x2 & 60.89 & 31 & 6 \\
3x3x4 & 62.91 & 33 & 12 \\
3x4x3 & 62.37 & 33 & 9 \\
5x3x2 & 71.70 & 31 & 10 \\
\hline
    \end{tabular}
}
\caption{Performance of the best 14 codes on Intel CascadeLake, AVX-512, with Clang15,  icc 19.1.3 and gcc 11.9. Only codes using 32 registers and 0 spill are considered here\label{fig:tables}. block size corresponds to $ui*uj+ui*uk+uk*uj$. $40$ means that the code requires $40$ zmm registers if all loads are scheduled before the computation. ILP corresponds here to $ui*uj$.}
\end{figure*}

\subsection{Impact of expected ILP}
Figure \ref{fig:cascadelakeilp} shows the performance according to the parallelism expected from unrolling, here $ui * uj$. Indeed, $uk$ does not create additionnal ILP. From this plot, we observe that for icc and clang, the best codes result from a high ILP. On the contrary, for gcc the code without spill with the highest expected ILP only reaches $42\%$ of the max performance. The maximum performance

\begin{figure*}[h!]
%  \begin{subfigure}[h]{0.45\textwidth}
%  \includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/icc-2021.2.0.pdf}
%  \caption{icc 2021.2.0}
%  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/icc-19.1.3p.pdf}
  \caption{Intel CascadeLake, AVX-512, icc 19.1.3}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/greedyp.pdf}
  \caption{Intel CascadeLake, AVX-512, Clang 15}
  \end{subfigure}\\
  \begin{subfigure}[h!]{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/cascadelake-64x256x64/gcc-11.9p.pdf}
  \caption{Intel CascadeLake, AVX-512, GCC 11.9}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/zen3-64x256x64/greedyp.pdf}
  \caption{AMD Zen 3, AVX2, Clang 15}
  \end{subfigure}\\
  \begin{subfigure}{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/zen3-64x256x64/greedy-zen2p.pdf}
  \caption{AMD Zen 2, AVX2, Clang 15}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}  
\includegraphics[width=\textwidth]{../benches/gemm/haswell-64x256x64/greedyp.pdf}
  \caption{Intel Haswell, AVX2, Clang 15}
  \end{subfigure}
  \caption{Block DGEMM Performance used on Intel CascadeLake according to expected ILP. Performance is given in \% of peak flops/cycle (here 32). The expected ILP is the product of the unrolled factors $ui * uj$. The block (BI,BK,BJ) is 64x256x64. Each code represents a different register block size, obtained through the three unrolling factors. Codes with register block size $\leq 32$ are in blue (should fit in registers whatever the schedule), other codes are in grey (with spill) and red (without spill).\label{fig:cascadelakeilp}}
\end{figure*}

\subsection{Correlation between compiler versions on Intel CascadeLake}

@@

The experiments show that the best performing codes have the same unrolling factors and the assembly code generated are the same. 
%\begin{figure*}[h!]
%  \caption{Block DGEMM Performance used on AVX2 architectures. Performance is given in \% of peak flops/cycle (here 8). The block (BI,BK,BJ) is 64x128x64. Each code represent a different register block size, obtained through the three unrolling factors. Codes with register block size $\leq 32$ are in blue (fit in register whatever the schedule), others in red. (c) show correlation between the two plots. $^*$ Performance scaled down to take into account turboboost effect, active on these machines\label{fig:avx2}}
%\end{figure*}
